First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2017-01-23 eg6-letters/DVCS-coh-He4/dvcs_he4_prl.tex
added the BH reaction and the BKM equations

A DVCS process on a nuclear target differ from single nucleon scattering in 
providing access to the measure two DVCS channels. In the coherent DVCS 
channel, the target nucleus ($A$) remains intact and recoils as a whole while 
emitting a real photon ($eA \rightarrow e' A' \gamma$), allowing to measure the 
nuclear GPDs of the target. In the incoherent channel, the nucleus breaks up 
and the DVCS takes place on a bound nucleon ($N$) that emits the final photon 
($eA \rightarrow e' N' \gamma$ X), enabling the GPDs measurement of the bound 
nucleons and to study the medium modifications of the nucleons in the nuclear 
medium via the GPDs.  Figure \ref{fig:diags} illustrates the dominant mechanism 
for the coherent DVCS channel on $^4$He. At sufficiently large squared electron 
momentum transfer $Q^2$ ($= -(k-k')^{2}$) and small squared momentum transfer 
$t$ ($= (p-p')^{2}$), the QCD factorization theorem predicts that the DVCS 
handbag diagram can be factorized into two parts, hard and soft parts 
\cite{Freund_Collins,Ji_Osborne}. The hard part includes photons-quark 
interaction and it is calculable through perturbative methods, while the 
soft/non-perturbative part is parametrized in terms of GPDs, which embed the 
partonic structure of the hadron.

@@

%DVCS selection
Identifying the coherent DVCS candidates is the first step of the data 
analysis. These are the events with one electron, one $^4$He, and at least one 
photon in the final state. Electrons were identified by passing the fiducial 
cuts and having signals in all the sub-detectors of CLAS spectrometer (drift 
chambers, Cherenkov counters, the standard CLAS electromagnetic calorimeter, 
and scintillators). $^4$He tracks were identified by passing all the 
geometrical, timing and quality cuts in the RTPC detector. The most energetic 
IC photon was considered as the DVCS photon candidate. Next, a 
$Q^{2}~>~[GeV^{2}/c^{2}]$ cut is applied on the DVCS candidates in order to 
ensure that the interaction occurs at the partonic level and the applicability
of the factorization in the DVCS handbag diagram. One the three final state 
particles were identified with their 3-momentum vectors, the exclusivity of the 
coherent DVCS events were ensured by applied a set of 3$\sigma$ exclusivity 
cuts. The exclusive cuts are: the co-planarity angle ($\Delta \phi$), missing 
energy, missing mass squared and missing transverse momentum in the 
$e'~^4He'~\gamma~X$ final state configuration, the missing mass squared in the 
$e'~^4He'~X$ and $e'~\gamma~X$ configurations, and finally the cone angle 
($\theta$) between the measured real photon and the missing particle in the 
$e'~^4He'~X$ configuration. Figure~\ref{fig:kin-cuts} presents four of the 
applied exclusivity cuts.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/coh_exc_cuts.pdf}
\caption{ The coherent DVCS exclusivity cuts. The black
distributions represent the coherent DVCS events candidate. The shaded
distributions represent the events which passed all the exclusivity cuts
except the quantity plotted. The vertical red lines represent $3\sigma$ cuts.}
\label{fig:kin-cuts}
\end{figure}

@@

%asymmetries

We identified several background contributions to the DVCS process, in 
particular accidental events and exclusive deeply virtual $\pi^0$ production 
(DV$\pi^0$P). The accidental
events where the different particles come from different events are 
suppressed by the limited phase space allowed by the exclusivity cuts. We 
estimate the accidental events to represent 4.1\% of our sample. This 
relatively large number is due to the small cross section of the DVCS. We 
evaluated this contribution by selecting events passing all our cuts but with 
particles originating from different vertices.

Another important source of background is the deeply virtual 
$\pi^0$ production (DV$\pi^0$P), which can easily be mistaken
with DVCS when one of the two photons of the $\pi^0$ decay is produced at
low energy in the laboratory frame. To estimate the importance of
this background, we developped an event generator 
that we calibrated to match our measured experimental yield of exclusive 
$pi^0$. We used this generator together 
with a GEANT3 simulation of our detection system to estimate the ratio 
of acceptance between DV$\pi^0$P where the two photons are detected and those
where only one photon is detected and would pass our 
DVCS selection cuts. This ratio obtained from simulation is then multiplied by 
the measured yield of DV$\pi^0$P events, indicating a contamination of 2 to 4\%. 
The study of systematics errors showed that the main contributions come from 
the choice of the DVCS exclusivity cuts (8\%) and the large binning size 
(5.1\%). However added quadratically, these errors sum up to 10\% which remain 
for all bins well below the statistical errors.

The beam-spin asymmetry is measurable using a polarized lepton beam on an 
unpolarized target (U). It is convenient to use the beam-spin asymmetry as a 
DVCS observable because most of the experimental normalization and acceptance 
issues cancel out in the asymmetry ratio. It is defined as in terms of the 
cross sections as:
  \begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
  where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the DVCS differential cross section for a positive (negative) beam helicity.

In terms of the collected number of events in each beam-helicity state 
($N^{+}$, $N^{-}$), A LU can be
expressed as:

\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} }.
\end{equation}
%
where $P_{B}$ is the beam polarization, and $N^{+}$ and $N^{-}$ are the number of events detected with positive and negative electron helicity, respectively. The statistical uncertainty of $A_{LU}$ is
%
\begin{equation}
   \sigma_{A_{LU}} = \frac{1}{P_{B}} \sqrt{ \frac{1 - (P_{B}A_{LU})^{2}}{N}}
\end{equation}
%
where $N (N^{+} + N^{-}) $ is the total number of measured events.
%

@@

%CFF
we extract the GPD $H$ in a model independent way (figure~\ref{fig:CFF_HA}).  
This is possible using the beam-spin asymmetry ($A_{LU}$) expression for 
spin-zero target at leading twist~\cite{BM_2009}:
\small
\begin{equation}
\begin{split}
A_{LU}(\phi) =~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\\
 \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( 
\Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}
\label{eq:A_LU-coh}
\end{split}
\end{equation}
\normalsize
%
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary and real parts of the CFF $\mathcal{H}_{A}$ associated to the GPD $H_A$. The $\alpha_{i}$'s are $\phi$-dependent kinematical factors that depend on the nuclear form factor $F_A$ and the independent variables $Q^2$, $x_{B}$ and $t$. These factors are simplified as:
%
\small
\begin{eqnarray}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi)\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi)\\ 
   \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right)\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2}
\end{eqnarray}
\normalsize
%

@@

\bibitem{CLAS_ref}
   B.A. Mecking {\it et al.}, Nucl.\ Inst.\ and Meth.\ A 503, 513 (2003).

>>>>

A DVCS process on a nuclear target differ from single nucleon scattering in 
providing access to the measure two DVCS channels. In the coherent DVCS 
channel, the target nucleus ($A$) remains intact and recoils as a whole while 
emitting a real photon ($eA \rightarrow e' A' \gamma$), allowing to measure the 
nuclear GPDs of the target. In the incoherent channel, the nucleus breaks up 
and the DVCS takes place on a bound nucleon ($N$) that emits the final photon 
($eA \rightarrow e' N' \gamma$ X), enabling the GPDs measurement of the bound 
nucleons and to study their modifications in the nuclear medium via the GPDs.  
Figure \ref{fig:diags} illustrates the dominant mechanism for the coherent DVCS 
channel on $^4$He. At sufficiently large squared electron momentum transfer 
$Q^2$ ($= -(k-k')^{2}$) and small squared momentum transfer $t$ ($= 
(p-p')^{2}$), the QCD factorization theorem predicts that the DVCS handbag 
diagram can be factorized into two parts, hard and soft parts 
\cite{Freund_Collins,Ji_Osborne}. The hard part includes photons-quark 
interaction and it is calculable through perturbative methods, while the 
soft/non-perturbative part is parametrized in terms of GPDs, which embed the 
partonic structure of the hadron. Experimentally, the DVCS reaction is 
indistinguishable from the Bethe-Heitler (BH) process, which is the reaction 
where the final photon is emitted either from the incoming or the outgoing 
leptons. The BH process is not sensitive to GPDs and does not carry information 
about the partonic structure of the hadronic target. The BH cross section is 
calculable from the well-known electromagnetic form factors.

@@

%DVCS selection
Identifying the coherent DVCS candidates is the first step of the data 
analysis. These events have one electron, one $^4$He, and at least one photon 
in the final state. Electrons were identified by passing the fiducial cuts and 
having signals in all the sub-detectors of CLAS spectrometer (drift chambers, 
Cherenkov counters, the standard CLAS electromagnetic calorimeter, and 
scintillators). $^4$He tracks were identified by passing all the geometrical, 
timing and quality cuts in the RTPC detector. The most energetic IC photon was 
considered as the DVCS photon candidate. Next, a $Q^{2}>1~[GeV^{2}/c^{2}]$ cut 
is applied on the DVCS candidates in order to ensure that the interaction 
occurs at the partonic level and the applicability of the factorization in the 
DVCS handbag diagram. Once the three final state particles were identified with 
their 3-momentum vectors, the exclusivity of the coherent DVCS events were 
ensured by applied a set of 3$\sigma$ exclusivity cuts. The exclusive cuts are: 
the co-planarity angle ($\Delta \phi$), missing energy, missing mass squared 
and missing transverse momentum in the $e'~^4He'~\gamma~X$ final state 
configuration, the missing mass squared in the $e'~^4He'~X$ and $e'~\gamma~X$ 
configurations, and finally the cone angle ($\theta$) between the measured real 
photon and the missing particle in the $e'~^4He'~X$ configuration.  
Figure~\ref{fig:kin-cuts} presents four of the applied exclusivity cuts.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/coh_exc_cuts.pdf}
\caption{Four of the seven coherent DVCS exclusivity cuts. The black 
distributions represent the coherent DVCS events candidate. The shaded 
distributions represent the events which passed all the exclusivity cuts except 
the quantity plotted. The vertical red lines represent $3\sigma$ cuts. The 
distributions from left to right and from top to bottoms are: $\Delta \phi$, 
missing energy and missing mass squared in $e'~^4He'~\gamma~X$, and the cone 
angle ($\theta$) between the measured and the calculated final state real 
photon.}
\label{fig:kin-cuts}
\end{figure}
After all the particle identification requirements on the individual particle 
of the coherent DVCS events and after all the exclusivity cuts, we ended up 
with about 3500 events. Figure \ref{fig:kin-coverage} presents the 
($Q^{2}$,$x_{B}$) and ($Q^{2}$,$-t$) kinematic coverage of the collected DVCS 
events.

@@

%asymmetries
Even with all the previously presented exclusive cuts, the selected events are
not all true DVCS events. We identified several background contributions to the 
coherent DVCS process, in particular accidental events and exclusive Deeply 
Virtual $\pi^0$ Production (DV$\pi^0$P). The accidental events where the 
different particles come from different events are suppressed by the limited 
phase space allowed by the exclusivity cuts. We estimate the accidental events 
to represent 4.1\% of our sample. This relatively large number is due to the 
small cross section of the DVCS. We evaluated this contribution by selecting 
events passing all our cuts but with particles originating from different 
vertices. Regarding the DV$\pi^0$P, which can easily be mistaken with DVCS when 
one of the two photons of the $\pi^0$ decay is produced at low energy in the 
laboratory frame. To estimate the importance of this background, we developed 
an event generator that we calibrated to match our measured experimental yield 
of exclusive $\pi^0$. We used this generator together with a GEANT3 simulation 
of our detection system to estimate the ratio of acceptance between DV$\pi^0$P 
where the two photons are detected and those where only one photon is detected 
and would pass our DVCS selection cuts. This ratio obtained from simulation is 
then multiplied by the measured yield of DV$\pi^0$P events, indicating a 
contamination of 2 to 4\%. The study of systematics errors showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%). However added quadratically, these errors sum up to 
10\% which remain for all bins well below the statistical errors.

The beam-spin asymmetry is our DVCS observable, which is measurable using a 
polarized lepton beam on an unpolarized target (U). It is convenient to use the 
beam-spin asymmetry as a DVCS observable because most of the experimental 
normalization and acceptance issues cancel out in the asymmetry ratio. It is 
defined in terms of the cross sections as:
  \begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
  where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the DVCS differential cross 
  section for a positive (negative) beam helicity. $A_{LU}$ can be simplified 
  in terms of the collected number of events in each beam-helicity state 
  ($N^{+}$, $N^{-}$) as:
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} }.
\end{equation}
where $P_{B}$ is the beam polarization, and $N^{+}$ and $N^{-}$ are the number 
of events detected with positive and negative electron helicity, respectively.  
The statistical uncertainty of $A_{LU}$ is
\begin{equation}
   \sigma_{A_{LU}} = \frac{1}{P_{B}} \sqrt{ \frac{1 - (P_{B}A_{LU})^{2}}{N}}
\end{equation}
where $N (N^{+} + N^{-}) $ is the total number of measured events.

%CFF
we extract the GPD $H$ in a model independent way (figure~\ref{fig:CFF_HA}).  
This is possible using the beam-spin asymmetry ($A_{LU}$) expression for 
spin-zero target at leading twist \cite{Kirchner:2003wt,Belitsky:2008bz}:
\small
\begin{equation}
\begin{split}
A_{LU}(\phi) =~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\\
 \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( 
\Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}
\label{eq:A_LU-coh}
\end{split}
\end{equation}
\normalsize
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the CFF $\mathcal{H}_{A}$ associated to the GPD $H_A$. The 
$\alpha_{i}$'s are $\phi$-dependent kinematical factors that depend on the 
nuclear form factor $F_A$ and the independent variables $Q^2$, $x_{B}$ and $t$.  
These factors are simplified as:
\small
\begin{eqnarray}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi)\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi)\\ 
   \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right)\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2}
\end{eqnarray}
\normalsize

@@

@@

\bibitem{CLAS_ref}
   B.A. Mecking {\it et al.}, Nucl.\ Inst.\ and Meth.\ A 503, 513 (2003)

\bibitem{Kirchner:2003wt} 
   A.~Kirchner and D.~Mueller, Eur.\ Phys.\ J.\ C {\bf 32}, 347 (2003)

\bibitem{Belitsky:2008bz} A.~V.~Belitsky and D.~Mueller,
     Phys.\ Rev.\ D {\bf 79}, 014017 (2009)
